---
title: "Untitled"
output: html_document
---

# Plotswith ggplot2

We are going to make plots in R using the ggplot2 package from the tidyverse.   

We will install these packages:

- tidyverse
- readxl
- here

```{r setup, message=FALSE}
library(tidyverse)  # install.packages("tidyverse")
library(readxl) #install.packages("readxl")
library(here)  # install.packages("here")
```



# California National Parks Data

```{r read-data}

ca_np <- read_csv(here("data/ca_np.csv"))
ci_np <- read_excel(here("data", "ci_np.xlsx"))
# also be written:
# ca_np <- read_csv("data/ca_np.csv")


```



### make my own first graph
#### different size header

We're going to make a line plot of year (x-var) vs. visitors (y-var) in ggplot
command + option/alt + i

```{r}
# trying ggplot2 with channel islands
ggplot(data = ci_np, aes(x= year, y = visitors)) + 
  geom_area()

# updata the geom to geom_point(), geom_col(), geom_area()
```

making a base layer for my channel islands plots:

```{r}
gg_base <- ggplot(data = ci_np, aes(x= year, y = visitors))
gg_base + geom_point()
```


### now let's start customizing
- line/point colors: color
- change line type: linetype
- change point style: pch
- change transparency: alpha

```{r}
gg_base + 
  geom_line(
    color = "purple",
    linetype = "dashed"
    
  )
```


activity: creat a ggplot scatterplot (geom_point) that has triangle symbols (hint: ?pch) of a non-default size (size = #) and that are an updated color.

```{r}
gg_base + 
  geom_point(
    color = "red", 
    pch = 2,
    size = 4
  )
```

### update graph aesthetics based on variables

```{r}
gg_base + 
  geom_point(
    aes(size = visitors, 
        color = year),
    pch = 10
  )
```

### use gg-themes for overall updates

- theme_light()
- theme_minimal()
- theme_bw()

```{r}
gg_base + 
  geom_area() +
  theme_light() 
```


### create mult-series line plot using data for all CA National Parks

```{r}
ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name)) +
  geom_line(aes(color = park_name))
```


```{r}
my_np_graph <- ggplot(data = ca_np, aes(x = year, y = visitors)) + 
  geom_point(aes(color = park_name), show.legend = FALSE) +
  facet_wrap(~park_name, scales = "free") +
  theme_bw()
```

export my graph as a .png
```{r}
ggsave(plot = my_np_graph, here("figures", "np_graph.png"))
```

